The algorithm aversion literature presents apparently contradictory findings: people sometimes reject algorithms that outperform humans, sometimes prefer them, and sometimes show no betrayal response to algorithmic failure. I argue these contradictions reflect a conceptual gap: the finance literature treats trust as a single dimension, conflating two structurally distinct modes. Drawing on philosophical trust theory (Baier, Holton, Hawley, Hardin), I distinguish \textit{evaluative trust} (evidence-conditional performance assessment) from \textit{delegative trust} (the grant of discretionary authority that is insensitive to ordinary performance signals but vulnerable to perceived betrayal). When delegative trust toward a category of trustee has never formed, I term this \textit{delegative foreclosure}: a structural closure that no amount of performance evidence is likely to reopen. I test this framework with two studies. First, AI-assisted classification of 9,000 CFPB complaint narratives reveals a cross-product delegative gradient --- mortgage complaints express significantly more delegative trust language than credit card complaints --- while evaluative trust remains flat, consistent with dimensional separability. Second, a pre-registered experiment ($N = 354$) finds that delegative trust is the dominant predictor of resistance to algorithm aversion ($\beta = -0.304$, $p < .001$; coefficient from aversion-coded dependent variable), while evaluative trust is nonsignificant. These results suggest that the contradictions in the literature arise from an unmeasured dimension: findings attributed to algorithm aversion or appreciation operate within evaluative trust, while the delegative dimension that governs advisory relationships has gone unrecognized.
