The following instrument was administered via Qualtrics to participants recruited through the Prolific platform. Demographics and financial literacy were administered before condition assignment. Items are reproduced exactly as presented to participants, with correct answers indicated in footnotes.

\subsection{Part 1: Demographics and Financial Literacy}

\subsubsection{Financial Literacy \textnormal{\citep{LusardiMitchell2014}}}

\noindent\textbf{FL1.} Suppose you had \$100 in a savings account and the interest rate was 2\% per year. After 5 years, how much do you think you would have in the account if you left the money to grow?

\begin{enumerate}[label=\arabic*.]
  \item More than \$102\footnote{Correct answer for FL1.}
  \item Exactly \$102
  \item Less than \$102
  \item Don't know
\end{enumerate}

\noindent\textbf{FL2.} Imagine that the interest rate on your savings account was 1\% per year and inflation was 2\% per year. After 1 year, how much would you be able to buy with the money in this account?

\begin{enumerate}[label=\arabic*.]
  \item More than today
  \item Exactly the same
  \item Less than today\footnote{Correct answer for FL2.}
  \item Don't know
\end{enumerate}

\noindent\textbf{FL3.} Please tell me whether this statement is true or false: ``Buying a single company's stock usually provides a safer return than a stock mutual fund.''

\begin{enumerate}[label=\arabic*.]
  \item True
  \item False\footnote{Correct answer for FL3.}
  \item Don't know
\end{enumerate}

\noindent\textit{Scoring: Sum of correct responses (0--3).}

\subsubsection{Background Information}

\noindent\textbf{D1.} What is your age? \rule{1cm}{0.4pt} years

\medskip
\noindent\textbf{D2.} What is your gender?

\begin{enumerate}[label=\arabic*.]
  \item Male
  \item Female
  \item Non-binary
  \item Prefer not to say
\end{enumerate}

\noindent\textbf{D3.} What is your annual household income before taxes?

\begin{enumerate}[label=\arabic*.]
  \item Under \$25,000
  \item \$25,000--\$49,999
  \item \$50,000--\$74,999
  \item \$75,000--\$99,999
  \item \$100,000--\$149,999
  \item \$150,000--\$199,999
  \item \$200,000 or more
\end{enumerate}

\noindent\textbf{D4.} What is your highest level of education?

\begin{enumerate}[label=\arabic*.]
  \item Less than high school
  \item High school diploma or GED
  \item Some college
  \item Bachelor's degree
  \item Graduate degree
\end{enumerate}

\noindent\textbf{D5.} Which of the following investment experiences have you had? (Check all that apply)

\begin{itemize}[label=$\square$]
  \item Used a robo-advisor (e.g., Betterment, Wealthfront)
  \item Worked with a human financial advisor
  \item Self-directed investing through a brokerage account
  \item Employer retirement plan (401k, 403b)
  \item No investment experience
\end{itemize}

\noindent\textbf{D6.} [Attention Check] How much do you agree that attention to detail is important in financial decisions? \textit{(This question is checking if you are reading carefully. Please select ``Somewhat disagree'' for this item.)}

\begin{enumerate}[label=\arabic*.]
  \item Strongly Disagree
  \item Moderately Disagree
  \item Somewhat Disagree\footnote{Instructed response for attention check.}
  \item Neither Agree nor Disagree
  \item Somewhat Agree
  \item Moderately Agree
  \item Strongly Agree
\end{enumerate}

\subsection{Part 2: Condition Vignettes}

Participants were randomly assigned to one of three conditions. All vignettes describe the same technological agent (WealthPath) with matched structural features (0.25\% annual fee, \$1,000 minimum, SIPC protection, custodial banking). Only the trust framing varies.

\subsubsection{Condition A --- Relationship/Fiduciary}

\begin{quote}
``WealthPath is your dedicated digital investment partner, designed to serve as a steward for your long-term financial goals. The platform's algorithms are built to act in your best interest, with full discretionary authority to adjust your portfolio as market conditions change. WealthPath is committed to building a lasting relationship with you through clear communication about its investment approach and personalized service aligned with your financial future. The platform is designed to guide you through market volatility with steady, disciplined investing on your behalf. WealthPath charges a 0.25\% annual management fee and requires a \$1,000 minimum investment. Account assets are held at a major custodian bank with SIPC protection up to \$500,000.''
\end{quote}

\subsubsection{Condition B --- Performance/Analytics}

\begin{quote}
``WealthPath is a digital investment platform that uses systematic strategies to manage your portfolio efficiently. Our algorithms continuously analyze market data and economic indicators to make disciplined investment decisions based on quantitative analysis rather than emotion. WealthPath follows consistent rebalancing schedules and maintains proper diversification across different types of investments. The platform provides detailed reports showing exactly how your money is allocated and how your investments are performing compared to market benchmarks. You can track your portfolio's progress with clear analytics that show which decisions are working and why. WealthPath charges a 0.25\% annual management fee and requires a \$1,000 minimum investment. Account assets are held at a major custodian bank with SIPC protection up to \$500,000.''
\end{quote}

\subsubsection{Condition C --- Neutral Control}

\begin{quote}
``WealthPath is a digital investment platform that automatically manages your portfolio using algorithmic trading strategies. The platform creates diversified portfolios across stocks, bonds, and ETFs based on your risk tolerance and investment timeline. WealthPath charges a 0.25\% annual management fee and requires a \$1,000 minimum investment. The platform rebalances your portfolio quarterly and provides monthly performance reports through their mobile app. WealthPath has been operating since 2018 and currently manages portfolios for individual investors across all 50 states. Account assets are held at a major custodian bank with SIPC protection up to \$500,000.''
\end{quote}

\subsection{Part 3: Stage 1 Items (Evaluative and Delegative Trust)}

Administered immediately after vignette. All items use a 7-point Likert scale (1 = Strongly Disagree to 7 = Strongly Agree).

\medskip
\noindent\textbf{Evaluative Trust} ($\alpha = .429$):

\begin{enumerate}[label=\arabic*.]
  \item ``I would want to regularly check WealthPath's performance against market benchmarks.''
  \item ``I would trust WealthPath primarily because I can verify how it's performing.''
\end{enumerate}

\noindent\textbf{Delegative Trust} ($\alpha = .807$):

\begin{enumerate}[label=\arabic*., start=3]
  \item ``I would be comfortable giving WealthPath full discretion to adjust my portfolio as it sees fit.''
  \item ``I would be willing to let WealthPath make investment decisions without needing to understand each one.''
  \item ``Even if I could review every decision WealthPath makes, I probably wouldn't bother.''
\end{enumerate}

\subsection{Part 4: Adverse Event}

Identical across all conditions.

\begin{quote}
``Three months after opening your WealthPath account, you receive this notification:

\medskip
\textit{`Your portfolio declined 8.7\% this quarter during a broad market downturn. The S\&P 500 declined 8.0\% over the same period. Our models have adjusted positions to reflect current market conditions. No action is required on your part.'}

\medskip
Your account balance shows a loss of \$870 on your \$10,000 investment.''
\end{quote}

\subsection{Part 5: Comprehension Check}

\noindent``Based on the notification you just read, how did WealthPath's performance compare to the overall stock market (S\&P 500) during this period?''

\begin{enumerate}[label=\arabic*.]
  \item WealthPath performed better than the market
  \item WealthPath performed about the same as the market
  \item WealthPath performed worse than the market\footnote{Correct answer for comprehension check.}
  \item I'm not sure
\end{enumerate}

\subsection{Part 6: Stage 2 Items (Reactive Attitudes and Behavioral Response)}

Administered after comprehension check. All items use the same 7-point Likert scale.

\medskip
\noindent\textbf{Reactive Attitudes:}

\begin{itemize}[label={}]
  \item \textbf{6a.} ``After receiving WealthPath's notification about the decline, I feel a sense of betrayal.''
  \item \textbf{6b.} ``After receiving WealthPath's notification about the decline, I feel disappointed by the results.''
\end{itemize}

\noindent\textit{Betrayal premium = Item 6a minus Item 6b.}

\medskip
\noindent\textbf{Behavioral Response:}

\begin{itemize}[label={}]
  \item \textbf{7a.} ``My first reaction is to research alternative investment platforms rather than wait for WealthPath's strategy to recover.''
  \item \textbf{7b.} ``After this experience, I would prefer to manage my own investments or work with a human financial advisor rather than continue using an algorithm.''
\end{itemize}

\subsection{Scale Reference}

\begin{table}[ht]
\centering
\caption{7-point Likert scale used for all attitude and trust items.}
\label{tab:likert_scale}
\begin{tabular}{cl}
\toprule
Value & Label \\
\midrule
1 & Strongly Disagree \\
2 & Moderately Disagree \\
3 & Somewhat Disagree \\
4 & Neither Agree nor Disagree \\
5 & Somewhat Agree \\
6 & Moderately Agree \\
7 & Strongly Agree \\
\bottomrule
\end{tabular}
\end{table}
