This paper introduced a two-dimensional framework for understanding trust in financial advisory relationships, distinguishing evaluative trust (evidence-conditional performance assessment) from delegative trust (the grant of discretionary authority with acceptance of opacity). Three empirical contributions support the framework's core claims.

First, dimensional separability is demonstrated across methods. In naturalistic CFPB complaint narratives, delegative trust indicators track product characteristics (relationship opacity and duration) while evaluative indicators remain flat --- a pattern inconsistent with unidimensional models. In the Prolific experiment, the two trust composites share only 6\% of their variance and predict non-overlapping sets of behavioral outcomes.

Second, delegative trust predicts resilience to both exit and algorithm aversion after adverse performance, while evaluative trust predicts neither. The H3 result --- that delegative trust is the dominant predictor of resistance to algorithm aversion --- reverses the pre-registered prediction. The reversal strengthens the evidence for dimensional separability: the two trust modes make divergent predictions about post-shock behavior, and it is the delegative dimension, not the evaluative one, that predicts who remains with the algorithm. The pre-registered expectation that evaluative trust would drive algorithm aversion cannot be assessed given the low reliability of the evaluative composite ($\alpha = .429$); what the data clearly show is that resistance to aversion comes from the delegative dimension.

Third, the framework resolves a decade-long apparent contradiction in the algorithmic decision-making literature. Algorithm appreciation \citep{LoggMinsonMoore2019} and algorithm aversion \citep{DietvorstSimmonsMassey2015} are both movements within evaluative trust under different signal conditions. The absence of betrayal aversion toward algorithms \citep{kormylo2026} reflects the other side of the framework: delegative trust does not form under default experimental conditions, so the betrayal response --- its diagnostic reactive attitude --- has no ground to stand on. The second dimension has been absent from the experimental literature because the standard paradigms do not create conditions for its formation.

The concept of delegative foreclosure, derived from the intersection of our framework with \citeauthor{hardin1993street}'s \citeyearpar{hardin1993street} street-level epistemology, identifies a mechanism the existing literature cannot name: the structural closure of a trust mode toward a category of trustee, distinct from the downward revision of a trust estimate. This concept explains both the persistence of algorithm aversion in the face of superior performance evidence and the failure of our own manipulation to shift trust modes through framing. Trust postures toward algorithmic advisors are learned through experience, not installed by institutional language. The robo-advisor experience predictor in the Prolific data ($B = 0.729$, $p = .001$) provides correlational support for this experiential-learning mechanism.

For practitioners, the framework inverts a common design assumption. The robo-advisory industry's emphasis on transparency, explainability, and performance dashboards is optimized for evaluative trust --- and may actively undermine the delegative trust that predicts long-term client retention. Platforms seeking to reduce algorithm aversion and exit should invest not in better performance communication but in the institutional conditions that permit delegative trust to form: fiduciary framing, discretionary authority structures, and sustained interaction that allows vulnerability to be experienced as safe.

For regulators, the transparency paradox formalized in Claim~3 poses a genuine di\-lemma. Explainability mandates that make algorithmic decisions more legible may shift users from delegative to evaluative trust equilibria, increasing sensitivity to short-term performance fluctuations and raising the probability of exit --- the opposite of the regulatory intent. This does not argue against transparency as a value, but it suggests that disclosure regimes should be designed with attention to their effects on trust mode, not merely on information availability.

The paper's limitations --- particularly the evaluative trust measurement problem and the manipulation failure --- constrain the strength of causal claims. But the convergence of naturalistic and experimental evidence, across different populations, products, and measurement approaches, establishes the evaluative/delegative distinction as a productive framework for understanding trust in financial advisory relationships. The faithful machine is not the one that performs best on benchmarks. It is the one that earns, and honors, the grant of discretionary authority.
