% ===================================================================
% DATA AND METHODS
% ===================================================================
% Converted from docs/intro_methods_prose_draft.md
% ===================================================================

This study employs two complementary empirical approaches: a naturalistic analysis of consumer complaint narratives from the CFPB database and a pre-registered between-subjects experiment conducted on the Prolific platform. The two analyses are methodologically independent --- different populations, product domains, measurement instruments, and analytic strategies --- and are designed to provide convergent evidence for the evaluative/delegative trust framework through triangulation rather than mutual dependence.

\subsection{CFPB Complaint Database Analysis}
\label{sec:methods_cfpb}

\subsubsection{Sample}

We sampled 9,000 consumer complaint narratives from the CFPB database, equally distributed across three product categories: mortgage ($n = 3{,}000$), checking and savings ($n = 3{,}000$), and credit card ($n = 3{,}000$). These categories were selected to vary along the dimension our framework identifies as theoretically relevant: the degree to which the consumer--institution relationship involves delegation of discretionary authority. Mortgage relationships are long-duration, high-stakes, and involve ongoing servicing decisions (escrow management, modification processing, loss mitigation) that consumers cannot easily monitor. Credit card relationships are largely transactional, with observable terms and assessable performance on standard metrics (interest rates, fees, dispute resolution). Checking and savings accounts fall between these poles. The sample was drawn from narratives with sufficient text for classification (minimum 50 words) filed between 2018 and 2024.

\subsubsection{Classification Procedure}

Each narrative was independently classified by two large language models --- Claude 3.5 Sonnet (Anthropic; \texttt{claude-3-5-sonnet-20241022}) and GPT-4o (OpenAI) --- using identical classification prompts. The prompts defined the evaluative and delegative trust constructs using language drawn from the theoretical framework and asked each model to provide three outputs for each narrative: (a)~a categorical classification (evaluative, delegative, or unclassifiable), (b)~continuous scores on each dimension (0--1 scale), and (c)~a confidence rating (0--1 scale). The dual-model approach serves as an automated inter-rater reliability check: agreement between independently developed language models, trained on different data and using different architectures, provides stronger evidence for classification validity than either model alone.

Of the 9,000 narratives, Claude 3.5 Sonnet produced valid classifications for 8,991 (99.9\%) and GPT-4o for 8,952 (99.5\%). The two failure sets are completely disjoint --- no narrative failed in both models --- so the intersection comprises $N = 9{,}000 - 9 - 48 = 8{,}943$ narratives successfully classified by both. All dual-model analyses use this $N = 8{,}943$ sample.

\subsubsection{Moral Foundations Validation}

As an independent construct validity check, all narratives were scored using the extended Moral Foundations Dictionary \citep[eMFD;][]{Hopp2021}, which provides continuous scores on five moral foundations: care, fairness, loyalty, authority, and sanctity. The theoretical prediction is that delegative trust violations should activate relational moral foundations --- particularly loyalty, the foundation most closely associated with betrayal and in-group allegiance --- at higher rates than evaluative trust complaints.

\subsubsection{Human Validation}

I independently classified a confidence-stratified subsample of 100 narratives to assess whether AI classifications capture distinctions recognizable to a human coder. Both models reported high confidence overall: Sonnet's mean confidence was 0.882 and GPT-4o's was 0.877, with 68.8\% of Sonnet classifications at confidence $\geq 0.90$. To ensure the human validation sample included difficult cases, the subsample was stratified by Sonnet confidence level (30 low-confidence, 40 medium, 30 high-confidence) at relatively generous thresholds rather than sampling proportionally from the confidence distribution.

\subsection{Prolific Experiment}
\label{sec:methods_prolific}

\subsubsection{Design}
\label{sec:methods_design}

The experiment employed a between-subjects design with three conditions. Participants were randomly assigned to encounter a hypothetical robo-advisory platform (WealthPath) under one of three framings: fiduciary/\allowbreak relationship (Condition~A), performance/\allowbreak analytics (Condition~B), or neutral control (Condition~C). All participants then experienced an identical adverse performance event before completing dependent measures. The study was pre-registered at AsPredicted (\#275,531).

\subsubsection{Participants}
\label{sec:methods_participants}

Demographics and the financial literacy battery were administered before condition assignment to prevent the vignette from priming financial literacy responses.

We recruited 471 participants through the Prolific platform, targeting U.S.-based adults with balanced demographic representation. After pre-registered exclusions, the final sample comprised $N = 354$ participants (Condition~A: $n = 115$; Condition~B: $n = 127$; Condition~C: $n = 112$). The exclusion funnel proceeded as follows: 471 initial completes $\to$ 468 (excluding survey previews by the author during instrument testing) $\to$ 451 (excluding pilot responses) $\to$ 446 (Prolific platform matching) $\to$ 357 (minimum completion time) $\to$ 354 (dual attention/comprehension check). Robustness analyses using stricter exclusion criteria (attention-only: $N = 342$; comprehension-only: $N = 289$) produce identical patterns of significance.

\subsubsection{Materials}
\label{sec:methods_materials}

\paragraph{Condition vignettes.} Each vignette described WealthPath using matched structural features (0.25\% annual fee, \$1,000 minimum, SIPC protection, custodial banking) while varying the trust framing. Condition~A emphasized fiduciary cues: the platform as a ``dedicated digital investment partner'' and ``steward'' with ``full discretionary authority'' and a commitment to ``building a lasting relationship.'' Condition~B emphasized evaluative cues: ``systematic strategies,'' ``quantitative analysis,'' ``detailed reports,'' benchmark comparison, and analytics showing ``which decisions are working and why.'' Condition~C provided a bare product description with no emphasis in either direction.

Critically, all three vignettes described the same technological agent --- no human advisors were introduced in any condition. This addresses a confound in earlier algorithm trust research where fiduciary framing was bundled with human agency.

\paragraph{Adverse event.} All participants received an identical notification: their WealthPath portfolio declined 8.7\% during a quarter in which the S\&P~500 declined 8.0\%, producing a loss of \$870 on a \$10,000 investment. The 70-basis-point underperformance was designed to be attributionally ambiguous --- enough to register as underperformance for attentive evaluators, but close enough to the benchmark to be dismissible by delegative trustors.

\subsubsection{Measures}
\label{sec:methods_measures}

All items used a 7-point Likert scale (1 = Strongly Disagree to 7 = Strongly Agree) unless otherwise noted.

\paragraph{Evaluative trust} (2 items, $\alpha = .429$). Items assessed performance-monitoring orientation: ``I would want to regularly check WealthPath's performance against market benchmarks'' and ``I would trust WealthPath primarily because I can verify how it's performing.'' The low reliability of this composite is a recognized limitation; the scale captures benchmark-monitoring orientation but fails to cohere as a unitary construct. We report results using the composite but note the attenuation of evaluative trust effects throughout.

\paragraph{Delegative trust} (3 items, $\alpha = .807$). Items assessed willingness to grant discretionary authority: ``I would be comfortable giving WealthPath full discretion to adjust my portfolio as it sees fit''; ``I would be willing to let WealthPath make investment decisions without needing to understand each one''; and ``Even if I could review every decision WealthPath makes, I probably wouldn't bother.''

\paragraph{Betrayal} (1 item). ``After receiving WealthPath's notification about the decline, I feel a sense of betrayal.''

\paragraph{Disappointment} (1 item). ``After receiving WealthPath's notification about the decline, I feel disappointed by the results.''

\paragraph{Betrayal premium.} Computed as the difference between betrayal and disappointment scores (Item~6a minus Item~6b), operationalizing \citeauthor{Holton1994}'s \citeyearpar{Holton1994} reactive attitude distinction.

\paragraph{Exit intent} (1 item). ``My first reaction is to research alternative investment platforms rather than wait for WealthPath's strategy to recover.''

\paragraph{Algorithm aversion} (1 item). ``After this experience, I would prefer to manage my own investments or work with a human financial advisor rather than continue using an algorithm.'' This item operationalizes \citeauthor{DietvorstSimmonsMassey2015}'s \citeyearpar{DietvorstSimmonsMassey2015} specific definition of algorithm aversion as the shift from algorithmic to human judgment, distinct from general exit-seeking behavior.

% FLAG: VERIFY AGAINST DATA — the 81.6% comprehension check pass rate needs to be confirmed against actual Prolific data
\paragraph{Comprehension check} (1 item, administered immediately after the adverse event). Participants were asked how WealthPath's performance compared to the S\&P~500, with four response options. The correct answer (WealthPath performed worse) was selected by 81.6\% of the final sample. Comprehension pass/fail was coded as a covariate in all analyses. Following pre-registered protocol, participants who failed the comprehension check were retained in the primary analyses, with robustness checks excluding them.

\paragraph{Financial literacy} (3 items). The \citet{LusardiMitchell2014} ``Big Three'' financial literacy battery: compound interest, inflation, and diversification items. Responses were coded as correct/incorrect and summed to create a 0--3 financial literacy composite.

\paragraph{Attention check} (1 item). A directed-choice item administered in the demographic battery prior to condition assignment, instructing participants to select a specific response option. The pre-registered exclusion rule uses a dual-failure criterion: participants were excluded from the primary sample only if they failed both the attention check and the comprehension check. Robustness analyses using a stricter attention-only exclusion rule ($N = 342$) and a comprehension-only rule ($N = 289$) produce identical patterns of significance.

\paragraph{Demographics.} Age, income, investment experience, education, and prior robo-advisor experience, coded as a binary variable derived from a multi-select investment experience item that also captured human advisor use, self-directed brokerage investing, and employer retirement plan participation.

\subsubsection{Analytic Strategy}
\label{sec:methods_analysis}

The pre-registered analysis plan specified five hypothesis tests:

H1 (Trust Mode Separation) was tested via one-way ANOVA comparing evaluative and delegative trust composites across conditions, with planned contrasts between Conditions~A and~B.

H2 (Betrayal Premium) was tested via one-way ANOVA comparing the betrayal premium (6a minus 6b) across conditions, with the prediction that Condition~A would produce the largest premium.

H3 (Algorithm Aversion Mechanism) was tested via OLS multiple regression predicting algorithm aversion (Item~7b) from evaluative trust, delegative trust, and condition dummies, with the pre-registered prediction that evaluative trust would be the significant predictor.

H4 (Delegative Resilience) was tested via OLS multiple regression predicting exit intent (Item~7a) from evaluative trust, delegative trust, and condition dummies.

H5 (Financial Literacy Interaction) was tested via moderated regression including a financial literacy $\times$ condition interaction term predicting delegative trust.

All analyses were conducted using the pre-registered exclusion criteria. Robustness analyses using stricter attention-only ($N = 342$) and comprehension-only ($N = 289$) exclusion thresholds are reported alongside primary results.
