The results reported above provide convergent evidence from two independent methods that evaluative and delegative trust are separable dimensions with distinct behavioral signatures. They also reveal important boundary conditions on the framework --- most notably, the failure to manipulate trust modes through vignette framing --- that are themselves theoretically informative. We organize the discussion around four themes: what the manipulation failure teaches, the pattern of dimensional separability across methods, the reconciliation of the algorithm aversion literature, and the practical implications of delegative foreclosure.

\subsection{What the Manipulation Failure Teaches}
\label{sec:disc_manipulation}

The most straightforward prediction of our framework --- that fiduciary framing would elevate delegative trust and performance framing would elevate evaluative trust (H1) --- was not supported. All between-condition comparisons were nonsignificant (all $p > .22$, all $\eta^2 < .01$). The manipulation did not shift participants between trust modes.

We do not treat this as evidence against the evaluative/delegative distinction. The distinction is supported by the individual-differences analyses (H3, H4) and by the CFPB naturalistic data, neither of which depends on the manipulation. But the failure requires explanation, and the explanation is substantive rather than merely methodological.

\citeauthor{hardin1993street}'s \citeyearpar{hardin1993street} street-level epistemology provides the key. If trust postures are learned Bayesian priors --- dispositions the trustor brings to the encounter based on accumulated experience with a category of interaction --- then a brief vignette describing a robo-advisor in fiduciary terms cannot install a delegative posture in participants who have not learned, through experience, that algorithmic systems can be trusted with discretionary authority. The manipulation asks participants to \emph{imagine} delegating judgment; it cannot provide the experiential grounding that delegative trust requires.

The robo-advisor experience predictor supports this interpretation directly. Prior experience with robo-advisory platforms was the strongest predictor of delegative trust in the sample ($B = 0.729$, $p = .001$), substantially stronger than any experimental manipulation. Participants who had actually used algorithmic advisors --- who had the experiential encounters \citeauthor{hardin1993street}'s framework identifies as necessary for trust formation --- exhibited higher delegative trust regardless of condition assignment. This is not a confound; it is the mechanism. Delegative trust toward algorithms forms through repeated encounters in which vulnerability is experienced as safe, not through institutional framing alone.

This finding has implications for both research and practice. Experimentally, it suggests that single-session vignette designs --- prevalent in the algorithm aversion literature\VERIFYSAGE{Confirm whether ``dominant paradigm'' is fair; cite any multi-session or longitudinal algorithm aversion studies if they exist.} --- may be structurally unable to study delegative trust formation, because they cannot provide the experiential history that delegative trust requires. Longitudinal designs, in which participants interact with algorithmic advisors over multiple sessions and encounter both favorable and adverse outcomes, may be necessary. For practitioners, the implication is that onboarding framing is insufficient: fiduciary language in terms of service or welcome screens is unlikely to build delegative trust without a sustained relationship that gives the framing experiential content.

\subsection{Dimensional Separability Across Methods}
\label{sec:disc_separability}

The core theoretical claim --- that evaluative and delegative trust are separable dimensions responding to different signal classes (Claim~1) --- is supported by two independent lines of evidence.

In the CFPB analysis, the cross-product delegative gradient provides the cleanest test. Delegative trust indicators decline monotonically from mortgage (21.1\%) to checking/savings (16.4\%) to credit card (12.3\%), tracking the theoretically predicted dimension of relationship opacity and duration. Evaluative trust indicators are flat across the same products (Sonnet $d = 0.030$ for mortgage vs.\ credit card). This asymmetric pattern --- one dimension varying while the other holds constant --- is the signature of dimensional separability. It is robust across two independent AI classifiers and survives control for narrative length (74.3\% of the gradient retained).

In the Prolific experiment, separability appears in the predictive structure. Delegative trust predicts both exit intent (H4: $B = -0.206$, $p = .001$) and algorithm appreciation --- that is, continued preference for algorithmic advice despite an adverse event (H3: $B = -0.326$, $p < .001$). Evaluative trust predicts neither (both $p > .6$). The two dimensions are not merely distinguishable in language --- they predict different behaviors in a controlled post-shock context. The correlation between the two trust composites ($r = .246$, 6\% shared variance) further supports separability: related but far from redundant.

The triangulation is strengthened by the methodological independence of these findings. The CFPB analysis uses naturalistic complaint text, AI-based classification, and a between-product design across banking products. The Prolific experiment uses self-report measures, a between-subjects vignette design, and an algorithmic investment context. The two analyses share no participants, no measurement instruments, no product domain, and no analytic strategy. That both converge on dimensional separability --- through different operationalizations, in different populations, across different financial products --- constitutes substantially stronger evidence than either alone.

\subsection{Reconciling Algorithm Aversion and Appreciation}
\label{sec:disc_reconciling}

The evaluative/\hspace{0pt}delegative framework resolves an apparent contradiction in the algorithmic decision-mak\-ing literature that has persisted for over a decade.

\citeauthor{LoggMinsonMoore2019}'s \citeyearpar{LoggMinsonMoore2019} algorithm appreciation and \citeauthor{DietvorstSimmonsMassey2015}'s \citeyearpar{DietvorstSimmonsMassey2015} algorithm aversion are not competing findings. Both document evaluative trust operating as our framework predicts: updating monotonically with performance signals. \citeauthor{LoggMinsonMoore2019}'s participants observed positive signals and increased trust; \citeauthor{DietvorstSimmonsMassey2015}'s participants observed negative signals and withdrew it. The disagreement in the literature arises from treating these as evidence about fundamentally different human--algorithm relationships, when they are movements along the same evaluative dimension under different evidentiary conditions.

\citet{kormylo2026} introduce a third finding that completes the picture: betrayal aversion reduces delegation to human experts but not to algorithmic experts. Their explanation --- that people do not experience betrayal toward machines --- is empirically robust. Our framework provides the structural account of \emph{why} the betrayal response is absent: betrayal is the reactive attitude specific to delegative trust violation, and delegative trust requires the attribution of discretionary authority to the trustee. Under default experimental conditions --- including \citeauthor{kormylo2026}'s trust games and delegation tasks --- participants approach algorithms as tools to be assessed, not agents to be entrusted with judgment. Delegative trust does not form, so the betrayal response has no ground to stand on.

The three findings thus decompose cleanly within the two-dimensional framework. \citet{LoggMinsonMoore2019} document evaluative trust responding to positive signals. \citet{DietvorstSimmonsMassey2015} document evaluative trust responding to negative signals. \citet{kormylo2026} document the absence of delegative trust under default conditions. The literature has been studying variation \emph{within} evaluative trust --- appreciating or averting based on observed performance --- without recognizing that a second dimension exists.

Our Prolific data add to this picture. The H3 result --- that delegative trust, not evaluative trust, is the dominant predictor of sustained algorithm appreciation after an adverse event --- demonstrates that the second dimension is not merely a theoretical possibility but an active predictor of behavior. Participants with high delegative trust did not switch to human advice after the performance shock, even though they experienced the same adverse event as participants with low delegative trust. The evaluative dimension, which the existing literature treats as the complete explanation for algorithm aversion, had no predictive power ($p = .614$). This is precisely what the framework predicts: algorithm aversion is an evaluative phenomenon, but sustained appreciation --- continued willingness to rely on algorithmic advice --- comes from the delegative dimension. Investors who have granted discretionary authority do not revert to performance-based evaluation after a single adverse signal.

This reinterpretation also generates a novel prediction that none of the three papers can produce independently: if experimental designs create conditions for delegative trust formation toward algorithms --- through fiduciary framing combined with sustained interaction --- then betrayal-like responses should emerge even toward algorithmic advisors. Our manipulation attempted this and failed, likely because a single vignette cannot provide the experiential grounding required. The prediction remains testable in longitudinal designs.

\subsection{Delegative Foreclosure and Design Implications}
\label{sec:disc_foreclosure}

The concept of delegative foreclosure --- introduced in the theoretical framework (Section~\ref{sec:theory}) as the closure of an entire trust mode toward a category of trustee --- connects the empirical findings to practical questions about robo-advisor design and financial regulation.

The persistence of algorithm aversion is a well-documented puzzle. \citet{DietvorstSimmonsMassey2015} find that even showing users evidence of algorithmic superiority does not fully overcome aversion. The standard explanation appeals to evaluative mechanisms: perhaps users weight errors more heavily than successes, or perhaps they hold algorithms to higher standards than human advisors. Our framework offers a structural alternative. For users who have foreclosed delegative trust toward algorithms, the performance evidence is addressing the wrong dimension. Their evaluative channel may be open --- they can acknowledge that the algorithm performs well on metrics --- but the delegative channel is closed. They cannot bring themselves to grant the algorithm discretionary authority, to accept the vulnerability that delegation requires. No amount of performance evidence is likely to reopen that channel, because the closure is structural, not evidentiary.

This analysis inverts conventional wisdom about robo-advisor design. The industry's dominant strategy --- increase transparency, provide performance dashboards, offer explainability tools --- is optimized for evaluative trust. It makes the algorithm's decisions legible, verifiable, and assessable against benchmarks. Under our framework, this approach succeeds for users already in evaluative mode \citep[who may appreciate the algorithm per][]{LoggMinsonMoore2019} but fails for users in delegative foreclosure (who need relational repair, not more data) and actively undermines nascent delegative trust in users who might otherwise develop it. This is the transparency paradox formalized in Claim~3: transparency builds evaluative trust ($\partial E/\partial \tau > 0$) but erodes the conditions under which delegative trust can form ($\partial \delta/\partial \tau < 0$).

The CFPB company-level analysis provides suggestive naturalistic evidence for this mechanism. Among mortgage companies with 50 or more complaints, the proportion of delegative complaints ranged from 3.6\% to 37.8\% --- a 34.2 percentage-point spread. While exploratory, this variation suggests that institutional behavior --- not just product type --- shapes which trust mode consumers activate. Companies whose servicing practices create conditions for experienced vulnerability (long hold times on modification decisions, opaque escrow management, inconsistent communication) may generate more delegative complaints precisely because they have first created and then violated the conditions for delegative trust. This is a hypothesis for future research, not a conclusion from the current data.

The practical implication for robo-advisory platforms is specific: if the goal is long-term client retention rather than initial adoption, platforms should invest in fiduciary framing, discretionary authority structures, and the institutional conditions that permit delegative trust to form over time. This means less emphasis on real-time portfolio dashboards and per-trade explanations, and more emphasis on the relationship-like features that signal the platform's acceptance of fiduciary responsibility. The platforms that have best retained assets under management may already be doing this intuitively; our framework provides the theoretical basis for what they are doing and why it works.

\subsection{Limitations}
\label{sec:disc_limitations}

Several limitations qualify the conclusions drawn above.

First, the evaluative trust composite in the Prolific experiment demonstrated poor internal reliability ($\alpha = .429$, two items). This attenuates any effects of evaluative trust on dependent variables and means that the null findings for evaluative trust in the H3 and H4 regressions cannot be interpreted as strong evidence of absence. The evaluative dimension may predict algorithm aversion and exit intent in ways our measurement was unable to detect. Future research should develop multi-item evaluative trust scales for algorithmic advisory contexts.

Second, the human validation of the CFPB classification involved a single coder (the author) rating 100 narratives. While categorical agreement with the AI models was fair ($\kappa = 0.264$--$0.361$) and improved dramatically at high confidence (93.3\% agreement), a single coder cannot establish inter-rater reliability in the conventional sense. The primary disagreement pattern --- the human coding more narratives as unclassifiable --- is systematic and interpretable, but future work should employ multiple independent human coders to better calibrate the AI classification system.

Third, the CFPB sample is censored toward consumers who have already experienced relationship breakdowns severe enough to motivate a formal complaint. The negative correlation between delegative and evaluative scores ($r = -.59$) likely reflects this censoring rather than a true inverse relationship between the constructs, as argued in the results section (Section~\ref{sec:results_cfpb}). But the censored sample also means we observe trust modes primarily in their violated state, which limits what the data can tell us about trust formation and maintenance.

Fourth, the robo-advisor experience predictor --- the strongest evidence for \citeauthor{hardin1993street}'s experiential-learn\-ing mechanism --- is correlational. We cannot rule out that pre-existing individual differences (risk tolerance, technology comfort, financial sophistication) drive both the decision to use robo-advisors and the tendency to report higher delegative trust, rather than the experience itself building delegative trust over time. Longitudinal designs that track trust formation through early interactions with algorithmic advisors would provide stronger evidence.

Fifth, the manipulation failure means that causal claims about the effect of institutional framing on trust mode are not supported by the experimental data. The individual-differences analyses (H3, H4) are associational: they show that delegative trust \emph{predicts} behavioral outcomes, not that it \emph{causes} them. The theoretical framework provides the causal architecture, and the CFPB naturalistic data provide ecological validity, but the experimental causal evidence that a fully pre-registered design was intended to provide is absent.

Finally, the CFPB database does not cover investment advisory services (SEC/FINRA jurisdiction), and the Prolific experiment uses hypothetical vignettes rather than real investment decisions with financial consequences. Neither study observes actual investor behavior in response to actual portfolio losses. The triangulation across methods partially addresses this concern --- naturalistic text and controlled experiment, banking products and algorithmic investment advice --- but the ultimate test of the framework requires observation of real advisory relationships through adverse market conditions.
