% ===================================================================
% INTRODUCTION
% ===================================================================
% Converted from docs/intro_methods_prose_draft.md
% ===================================================================

Why do some investors abandon algorithmic financial advisors after a single bad quarter while others remain patient through prolonged downturns? The behavioral finance literature offers two competing answers. \citet{DietvorstSimmonsMassey2015} document ``algorithm aversion'' --- the tendency to abandon algorithms after observing errors, even when the algorithm outperforms human alternatives. \citet{LoggMinsonMoore2019} document the opposite: ``algorithm appreciation,'' a systematic preference for algorithmic over human judgment in predictive domains. These findings have been treated as context-dependent results requiring reconciliation through moderator variables --- task type, outcome observability, individual differences (for reviews, see \citealt{BurtonSteinJensen2020}; \citealt{MahmudIslamEtAl2022}; \citealt{JussupowBenbasatHeinzl2024}) --- without a unifying theoretical account of why the same class of decision-aid produces opposite reactions.

We propose that the contradiction is an artifact of dimensional collapse. The finance literature treats trust as varying in degree but not in kind: investors trust more or less, along a single continuum. Drawing on philosophical accounts of trust that have not previously been applied to financial relationships --- particularly the work of \citet{Baier1986}, \citet{Holton1994}, \citet{Nickel2012}, and \citet{Hawley2014} --- we distinguish two qualitatively different trust modes. \textit{Evaluative trust} is evidence-conditional performance assessment: the investor monitors outcomes, compares them to benchmarks, and maintains the relationship insofar as the evidence supports it. Withdrawal takes the form of proportional disappointment. \textit{Delegative trust} is the grant of discretionary authority with acceptance of opacity: the investor transfers judgment to the advisor and refrains from continuous verification. Violation produces not disappointment but betrayal --- a reactive attitude whose intensity reflects the depth of the relationship rather than the magnitude of the performance shortfall.

This distinction resolves the algorithm aversion puzzle. \citeauthor{LoggMinsonMoore2019}'s appreciation and \citeauthor{DietvorstSimmonsMassey2015}'s aversion are both evaluative trust responding to performance signals --- positive in one case, negative in the other. They are movements along the same dimension under different evidentiary conditions, not contradictory findings about human--algorithm interaction. \citet{kormylo2026} introduce a third finding that completes the picture: betrayal aversion reduces delegation to human experts but not to algorithmic experts. Our framework explains why --- betrayal is the reactive attitude specific to delegative trust violation, and under default experimental conditions, delegative trust toward algorithms does not form. The literature has been studying variation within evaluative trust without recognizing that a second dimension exists.

We test the evaluative/delegative framework using two complementary empirical approaches. First, we analyze 9,000 consumer complaint narratives from the Consumer Financial Protection Bureau (CFPB) database, classifying them for trust mode using dual-model AI classification (Claude 3.5 Sonnet and GPT-4o). The central finding is a cross-product delegative gradient: delegative trust indicators decline monotonically from mortgage to checking/savings to credit card complaints, tracking the theoretically predicted dimension of relationship opacity and duration, while evaluative trust indicators remain flat. This asymmetric pattern --- one dimension varying while the other holds constant across product types --- is the signature of dimensional separability.

Second, we conduct a pre-registered between-subjects experiment ($N = 354$ after pre-registered exclusions) in which participants encounter a robo-advisory platform under fiduciary, performance, or neutral framing before experiencing an identical adverse performance event. The manipulation does not shift trust modes --- a theoretically informative failure we explain through \citeauthor{hardin1993street}'s \citeyearpar{hardin1993street} street-level epistemology of trust, which holds that trust postures are learned through experience rather than installed by framing. However, the individual-differences analyses reveal the framework's predictive power: delegative trust is the dominant negative predictor of algorithm aversion ($B = -0.326$, $p < .001$) and exit intent ($B = -0.206$, $p = .001$), indicating that higher delegative trust is associated with lower aversion and lower exit-seeking. Evaluative trust predicts neither outcome.

The framework reframes what it means for an algorithmic advisor to be trustworthy. The ``faithful machine'' is not the one that maximizes benchmark-adjusted returns; it is the one whose institutional design permits the formation and honoring of delegative trust --- the grant of discretionary authority that defines a fiduciary relationship.

Existing reviews have explicitly called for theory integration to resolve these contradictions \citep{BurtonSteinJensen2020, JussupowBenbasatHeinzl2024}; this paper answers that call. The paper makes three contributions. First, it introduces the evaluative/delegative distinction to behavioral finance, grounding it in philosophical trust theory and operationalizing it for empirical measurement. Second, it provides convergent evidence from naturalistic and experimental methods that the two trust modes are separable, respond to different signal classes, and predict distinct post-shock behaviors. Third, it resolves the algorithm aversion/appreciation contradiction by showing that both phenomena operate within the evaluative dimension, while the delegative dimension --- absent from the experimental literature because standard paradigms do not create conditions for its formation --- predicts which investors resist the aversion response. We introduce the concept of \textit{delegative foreclosure} --- the structural closure of a trust mode toward a category of trustee, derived from \citeauthor{hardin1993street}'s insight about self-reinforcing distrust --- to explain both the persistence of algorithm aversion in the face of superior performance evidence and the failure of institutional framing to shift trust postures.

The remainder of the paper proceeds as follows. Section~\ref{sec:theory} develops the theoretical framework, distinguishing evaluative and delegative trust, connecting them to the principal--agent literature and the Money Doctors model, and deriving five pre-registered hypotheses. Section~\ref{sec:data_methods} describes the data and methods for both empirical analyses. Section~\ref{sec:results_cfpb} reports the CFPB results. Section~\ref{sec:results_prolific} reports the Prolific experimental results. Section~\ref{sec:discussion} discusses implications, including the reconciliation of the algorithm aversion literature and the design implications of delegative foreclosure. Section~\ref{sec:conclusion} concludes.
